---
title: "alli_data"
author: "Ciera Martinez"
date: "July 11, 2017"
output: pdf_document
---

**Purpose**: The purpose is to to correlate compounds to oviposition index. Mike wants to use multivariable regression or decision/regression to predict behavior. He also said to use linear or nonparametric. 


```{r}
## Library Prereqs
library(lme4)
library(ggplot2)
library(languageR)
library(xlsx)
library(reshape2)

## Functions

## modelcheck <- function(model, h = 8, w = 10.5) { ## because plot(lmer.obj) doesn't work rs <- residuals(model) fv <- fitted(model) quartz(h = h, w = w) plot(rs ~ fv) quartz(h = h, w = w) plot(sqrt(abs(rs)) ~ fv) quartz(h = h, w = w) qqnorm(rs) qqline(rs) }  
```

### Read in Data

From .xls file.

```{r}

sheet1 <- read.xlsx("../data/raw/alli_example_data.xlsx", 
                sheetIndex = 1)

## Check Sheet 1
head(sheet1)
dim(sheet1)
duplicated(sheet1)

### These are averaged and normalized

sheet2 <- read.xlsx("../data/raw/alli_example_data.xlsx", 
                sheetIndex = 2)

## Check Sheet 2
head(sheet2)
dim(sheet2)
colnames(sheet2)

sheet3 <- read.xlsx("../data/raw/alli_example_data.xlsx", 
                sheetIndex = 3)

## Check Sheet 3
head(sheet3)
dim(sheet3)
```

## Cleaning, transforming, and merging

There were two different experiments that were performed.
1. Measuring of the yeast chemical compounds present, as detailed in `sheet1`
2. Measurments of fly behavior laying eggs with presense of said yeast `sheet2` and `sheet3`

I believe to do multivariate analysis on trying to find correlations between chemical compunds present and Oviposition.index, we just have to assume it is one experiment i.e. merge the data together.

Each measurement will also have additional columns corresponding to what is measured for each yeast type. 

```{r}
#########################
## Sheet 1
##########################

## What are these variance, range, mean columns? These don't make sense to me.
##        I am going to remove.

# Peak position doesn't really make sense either (prob describes chemical component peak?), but is we are just assuming it is the chemical compound called, I will just use that.
# Update: Looks like a peak can call an identical compound and this causes problems later
#         therefore I will merge this column with 

## Check
colnames(sheet1)

# Remove
sheet1 <- sheet1[,-c(9:11)]

# add new column
sheet1$compound_peak <- paste(sheet1$compound, sheet1$peak_position, sep = "_")

## Check
head(sheet1)

## Now get rid of 1 and 2
sheet1 <- sheet1[,-c(1,2)]

# Transform Sheet1 first melt the Yeast columns from wide to long
sheet1_melt <- melt(sheet1, id.vars = "compound_peak",
     variable.name = "yeast")

## Check
head(sheet1_melt)

# Transform Sheet 1 compound components from long to wid
sheet1_cast <- dcast(sheet1_melt, yeast ~ compound_peak, value.var = "value")

## [x] Something wrong. Is there duplicate rows? Yes. See above.

## Check
head(sheet1_cast)
colnames(sheet1_cast) #There are 35 compounds measured for each yeast

## rename back to original

sheet1 <- sheet1_cast

## Get rid of Yeast 3:6

sheet1 <- sheet1[-c(3:6),]
head(sheet1)
```

### Sheet 2 and 3 

```{r}
######################
## sheet2 and sheet3
######################

# make sure they have same columns
colnames(sheet3)
colnames(sheet2)

# add together
sheet23 <- rbind(sheet2, sheet3)

## Now merge all three by yeast
## First make sure they are named the same

colnames(sheet23)
summary(sheet23$Yeast)
summary(sheet1$yeast)

## Fix
colnames(sheet23)[2] <- "yeast"

sheet23$yeast <- gsub("Yeast 1", "Yeast.1", sheet23$yeast)
sheet23$yeast <- gsub("Yeast 2", "Yeast.2", sheet23$yeast)

## Now we can merge

## Check
dim(sheet23)
dim(sheet1)

## Merge
sheet123 <- merge(sheet23, sheet1, by = "yeast")

##Check
dim(sheet123)
head(sheet123)


write.csv(sheet123, "../data/output/combinedData.csv")
```



