---
title: "alli_data"
author: "Ciera Martinez"
date: "July 17, 2017"
output: pdf_document
---

**Purpose**: The purpose is to to correlate compounds to oviposition index. Mike wants to use multivariable regression or decision/regression to predict behavior. He also said to use linear or nonparametric. 

**About Data**

Three excel files -
`egglay..` - all the behavior data for all fly lines and all yeast. Trials are organized by different yeast (one yeast per sheet)

`gcms_vals_norepeats` - the first column here is the peak_position and we can use as a "name" for each peak. these are true values for reach averaged yeast

`gcms_vals_normalized_norepeats` - this is the data normalized (across rows). Basically the same as what I gave you before without the mean/variance columns and only the peak_position column
Ready to model when you are!

```{r}
## Library Prereqs
library(lme4)
library(ggplot2)
library(languageR)
library(xlsx)
library(reshape2)

## Functions

## modelcheck <- function(model, h = 8, w = 10.5) { ## because plot(lmer.obj) doesn't work rs <- residuals(model) fv <- fitted(model) quartz(h = h, w = w) plot(rs ~ fv) quartz(h = h, w = w) plot(sqrt(abs(rs)) ~ fv) quartz(h = h, w = w) qqnorm(rs) qqline(rs) }  
```

### Read in Data

From .xls file.

```{r}
############## egglay_liqagar_72hr_rawdata_formodel.xlsx
############## EL_RME1027
EL_RME1027 <- read.xlsx("../data/raw/egglay_liqagar_72hr_rawdata_formodel.xlsx", 
                sheetIndex = 1)

## Check 
head(EL_RME1027)
dim(EL_RME1027)
duplicated(EL_RME1027)

############## EL_RME1028
EL_RME1028 <- read.xlsx("../data/raw/egglay_liqagar_72hr_rawdata_formodel.xlsx", 
                sheetIndex = 2)

head(EL_RME1028)
dim(EL_RME1028)
duplicated(EL_RME1028)

############## EL_RME1512
EL_RME1512 <- read.xlsx("../data/raw/egglay_liqagar_72hr_rawdata_formodel.xlsx", 
                sheetIndex = 3)

## Check 
head(EL_RME1512)
dim(EL_RME1512)
duplicated(EL_RME1512)

############## EL_RME1511

EL_RME1511 <- read.xlsx("../data/raw/egglay_liqagar_72hr_rawdata_formodel.xlsx", 
                sheetIndex = 4)

head(EL_RME1511)
dim(EL_RME1511)
duplicated(EL_RME1511)

########## EL_RME1553
EL_RME1553 <- read.xlsx("../data/raw/egglay_liqagar_72hr_rawdata_formodel.xlsx", 
                sheetIndex = 5)

head(EL_RME1553)
dim(EL_RME1553)
duplicated(EL_RME1553)

############## gcms_vals_norepeats.xlsx
############## gcms_no_repeats
gcms_no_repeats <- read.xlsx("../data/raw/gcms_vals_norepeats.xlsx", 
                sheetIndex = 1)

## Check 
head(gcms_no_repeats)
dim(gcms_no_repeats)
duplicated(gcms_no_repeats)

############## gcms_vals_normalized_norepeats.xlsx
############## gcms_norm
gcms_norm <- read.xlsx("../data/raw/gcms_vals_normalized_norepeats.xlsx", 
                sheetIndex = 1)

## Check 
head(gcms_norm)
dim(gcms_norm)
duplicated(gcms_norm)
```


## Cleaning, transforming, and merging

There were two different experiments that were performed.
1. Measuring of the yeast chemical compounds present, as detailed in `../data/raw/gcms_vals_normalized_norepeats.xlsx`
2. Measurments of fly behavior laying eggs with presense of said yeast `egglay_liqagar_72hr_rawdata_formodel.xlsx`

I believe to do multivariate analysis on trying to find correlations between chemical compunds present and Oviposition.index, we just have to assume it is one experiment i.e. merge the data together.

```{r}
# Add all EL sheets

egg_lay <- rbind(EL_RME1553, EL_RME1511, EL_RME1512, EL_RME1028, EL_RME1027)

head(egg_lay)
dim(egg_lay)

summary(egg_lay)

#make date column be date
egg_lay$Date <- as.Date(egg_lay$Date , "%y-%d-%m")
```

## GCMS

```{r}
head(gcms_norm)

# make peak a factor
gcms_norm$peak_pos <- as.factor(gcms_norm$peak_pos)

# Transform - first melt the Yeast columns from wide to long
gcms_melt <- melt(gcms_norm, id.vars = "peak_pos",
     variable.name = "Yeast")

# Transform - compound components from long to wid
gcms_cast <- dcast(gcms_melt, yeast ~ peak_pos, value.var = "value")

# Check
dim(gcms_cast)

# rename back to original
gcms_norm <- gcms_cast

# Check and fix for Merge 
colnames(gcms_norm)[1] <- "Yeast"
colnames(egg_lay)

dim(gcms_norm)
dim(egg_lay)

# Merge
all_data <- merge(egg_lay, gcms_norm, by = "Yeast")

## check
head(all_data)
dim(all_data)

write.csv(all_data, "../data/output/combinedData_17July2017.csv")

